A deeper look into bug fixes: Patterns, replacements, deletions, and additions

### 2016-02-18 MSRChallenge2016.tex
Merged my part and Pan's, removed a lot of text

\begin{table}[!htb]
	\caption{Frequency of Bug Fixing Patterns}\label{tab:freqpattern}
	\centering
	\begin{tabular}{lrr} 
		\hline
		& \textbf{\#Appearance} & \textbf{Percentage}\\
		\hline
		%Buggy Files & 46,301,429 \\ 
		AMP Pattern & 901,083 & 1.95\%\\
		MSM Pattern & 783,073 & 1.69\%\\
		COM Pattern & 270,128 & 0.58\%\\ 
		CBC Pattern & 1,959,377 & 4.23\%\\  
		IAO Pattern & 1,308,006 & 2.82\%\\  
		ANC Pattern & 1,340,561 & 2.90\%\\  
		AOB Pattern & 128,016 & 0.28\%\\  
		COB Pattern & 262,915 & 0.57\%\\   
		\hline
	\end{tabular}
\end{table}

@@

\section{Statement-level mutations}

\texttt{Assert}	&-&7.48&3.76&0.53&8.30&23.05&0.31&20.04&4.90&4.62&1.30&13.50&7.23&0.03&4.95
\\ \hline
\texttt{Break} &1.00&-&4.08&0.60&9.93&26.03&0.13&25.39&2.48&1.57&1.79&8.39&11.73&0.10&6.77
\\ \hline
\texttt{Continue} &1.74&9.42&-&1.28&11.39&18.25&0.35&22.60&3.80&2.85&2.17&8.98&9.42&0.11&7.63
\\   \hline
\texttt{Do} &0.81&5.26&6.60&0.00&9.44&14.21&0.18&15.86&3.73&1.67&1.97&5.88&6.39&0.03&27.98
\\ \hline
\texttt{For} &0.86&6.28&3.19&0.79&-&22.89&0.09&21.08&5.01&3.34&1.87&10.01&10.71&0.08&13.79
\\ \hline
\texttt{If}&1.64&8.43&2.87&0.60&13.49&-&0.24&26.46&7.45&4.80&2.85&9.89&15.11&0.08&6.11
\\ \hline
\texttt{Label} &1.30&8.33&7.86&1.11&5.18&22.85&-&15.17&3.05&2.04&14.62&10.45&4.16&0.09&3.79
\\ \hline
\texttt{Return} &1.13&9.41&3.11&0.49&13.33&27.24&0.24&-&5.59&3.65&2.55&14.91&12.61&0.12&5.61
\\ \hline
\texttt{Case} &0.78&2.84&2.84&0.39&10.27&31.79&0.16&22.40&-&0.46&2.07&7.37&11.69&0.08&6.87
\\ \hline
\texttt{Switch}&1.14&2.72&3.80&0.55&11.07&34.14&0.13&21.86&0.75&-&1.53&8.65&9.02&0.05&4.58
\\ \hline
\texttt{Synch} &0.80&6.57&2.28&0.43&10.21&24.18&0.05&19.77&6.35&2.07&-&9.16&12.16&0.04&5.93
\\ \hline
\texttt{Throw} &2.11&6.57&2.58&0.48&11.87&18.84&0.17&32.28&4.64&3.30&2.74&-&10.08&0.07&4.27
\\ \hline
\texttt{Try}&0.71&7.41&3.02&0.66&11.73&27.75&0.11&23.24&5.63&2.65&2.58&8.99&-&0.09&5.42
\\ \hline
\texttt{TypeDecl} &0.00&4.51&7.52&1.00&10.28&21.05&0.50&17.79&6.02&1.75&2.01&9.27&11.53&-&6.77
\\ \hline
\texttt{While}& 0.72&8.02&3.82&1.96&23.16&19.78&0.12&16.48&6.56&3.09&1.64&6.81&7.80&0.04&-
			\\ \hline
		\end{tabular}
		}
		\caption{Likeliness to replace the statement in the first column for the statement in the top row, in Java}
\end{table*}

Most automatic program repair approaches mutate source code using predefined
mutation operators, which represent different ways of modifying source code. In
order to generate patches that make more sense to human developers, it would be
beneficial to first investigate how developers mutate source code to fix bugs. 
Thus, beyond characterizing the degree to which previously-proposed Java-specific
operators apply in our dataset, we further characterized the nature of
coarser-grained replacement operators, which are also commonly translated from C
techniques to Java.  We did this via a query that visits all fixing revisions in
the repository; as a result, we analyze files to which changes have been made
and thuswere already committed and are now being reviewed to fix an error.

\subsection{Replacements}
\todo{The biggest problem right now is the lack of integration between this
  study and Chu Pan's.  Please work together to find a way to unify it.  One of
  you will need to cut some of your text.  I don't care who, but we need to find
  a way to tell one story.}

Because direct diffs are difficult to identify, we heuristically approximate the
extent to which one statement type appears to be ``replaced'' by another across
this dataset.  To do this, for each modified file, we count the number of
appearances of each statement type in the file pre- and post-commit.  We then
compare the results to see how many of each statement type was removed, and how
many inserted, to roughly characterize the types of replacement that happen at a
per-file level.  For each statement type (as tagged by the Boa infrastructure
for Java), we process the results as follows: 
if number occurrences of one statement type decreased post-fix and the number of
another type increased, we say that
the first statement type was replaced by the second statement type for that
file.  We collate these results over all statement types and all files modified
in bug fixing commits in the dataset.

@@

The numbers in the slots of Table 1, represent the percentage rate at which the
elements of the leftmost column were replaced by the elements of the top row. In
the first column we have the list of statements that are being replaced (faulty
statements that were removed), and on the top row we have the list of statements
for which that statement was replaced for (the new statement put instead). For
example, if we take the intersection of row 5 (DoStatement row) and column 2
(AsserStatement column) of the data, we have a 0.81, which means that if we take
all the instances in which a Do statement was replaced by something else, the
amount of times that it was replaced by an Assert statement, represent a
0.81% of the times.

This study is directly applicable to automatic software repair. Most common
tools for automatic software repair~\cite{kim2013,weimer2009,legoues2012}
first find a faulty statement using several different fault localization
techniques~\cite{fry2010}, and once they have that faulty statement, they apply
different kinds of edits to it, usually at random. Which makes it equally likely
to replace an statement for another without looking at the kinds of statements
being replaced.

With this study, we are able to make a more accurate guess, when having a known
faulty statement, to which statement it is more likely to replace it for. By
taking a look at the data of Table 1, we can find the most likely replacements
that will end up being a successful patch.

For example, if we are fixing a Java program, and we have a faulty statement
that happens to be a For Statement. Then we can look up in Table 1 the For
Statement Row (6th row), and take a look at the probability of that statement to
be replaced by other kinds of statement. In this example we see that the top 3
most likely statements to replace a For Statement are an If statement with a
22.89\% likeliness, followed by a return statement with a 21.08\%, and third is
a While statement with a 13.79\% of likeliness to replace a For Statement. This
way we know that it is more likely for a programmer to replace a For Statement
for those three different kinds of statements and this provides guidelines to
the automatic bug repair approaches to consider this on their effort to find a
patch.

@@

\subsubsection{Least common replacements}
On the opposite side of the data, we can see that the least common replacement
was an Assert statement replacing a Type Declaration statement. This was the
only replacement to have zero appearances in our search through the 46,301,429
files being fixed. Close to it, as the second least common replacement was
removing a Do statement and insert a Type Declaration statement, with one
appearance in our code database, and on third place, removing a Label and insert
a Type Declaration Statement had one appearance in our code database.

@@

\begin{figure*}[!t] \centering
	\includegraphics[width=0.95\textwidth]{StmtType} \caption{For each kind
	of statement, how many bug fixing revisions add/delete statements of the
same kind?} \label{fig:StmtType} \end{figure*}

For each bug fixing revision $r$ and each statement kind $k$, we count how many
statements of kind $k$ are there in revision $r$ and $r-1$ ($r-1$ means the
revision before revision $r$), and then compare the two numbers. For example, if
the number of if statements in revision $r$ is more than the number in revision
$r-1$, we say bug fixing revision $r$ introduce new if statements. Note that
``introducing new if statements'' does not necessarily mean human developers
introduce them to fix bugs. However, our hope is that this analysis could still
provide us some hints, given the large number of bug fixing revisions we are
analyzing.

@@

For each bug fixing revision, the results discussed above consider both newly
added files and existing files. Since most automatic program repair approaches
do not generate new files to fix bugs, the rest of this section only considers
files that were already in the repositories when developers fixed bugs. Our
analysis shows that there are 46,301,429 of them. In these files, there is a
total of 517,281 statement kinds being replaced, and 511,198 replacing these
statements kinds.

We analyzed each of these files looking for statement kinds being either
replaced by other kinds, deleted or appended to the file.

We found that the most common Replacer statement kind, meaning the statement
kind that most commonly replaces others is the \textbf{If Statement} with
101,366 appearances. The least common Replacer, meaning the statement kind which
is less likely to replace other statement kinds is the \textbf{Type Declaration
  Statement} with 447 appearances.

The most common Replacee, which is the most likely statement kind to be replaced
by another statement kind is the \textbf{Return Statement} with 111,938
appearances. The least common Replacee, which is the least common statement to
be replaced by another statement kind is the \textbf{Type Declaration Statement}
with 399 appearances.

>>>>

\begin{table}[!htb]
	\centering
	\begin{tabular}{lrr} 
		\hline
		& \textbf{\#Appearance} & \textbf{Percentage}\\
		\hline
		%Buggy Files & 46,301,429 \\ 
		AMP Pattern & 901,083 & 1.95\%\\
		MSM Pattern & 783,073 & 1.69\%\\
		COM Pattern & 270,128 & 0.58\%\\ 
		CBC Pattern & 1,959,377 & 4.23\%\\  
		IAO Pattern & 1,308,006 & 2.82\%\\  
		ANC Pattern & 1,340,561 & 2.90\%\\  
		AOB Pattern & 128,016 & 0.28\%\\  
		COB Pattern & 262,915 & 0.57\%\\   
		\hline
	\end{tabular}
	\caption{Frequency of Bug Fixing Patterns}\label{tab:freqpattern}

\end{table}

\texttt{Assert}	&-&7.48&3.76&0.53&8.30&23.05&0.31&20.04&4.90&4.62&1.30&13.50&7.23&0.03&4.95
\\ \hline
\texttt{Break} &1.00&-&4.08&0.60&9.93&26.03&0.13&25.39&2.48&1.57&1.79&8.39&11.73&0.10&6.77
\\ \hline
\texttt{Continue} &1.74&9.42&-&1.28&11.39&18.25&0.35&22.60&3.80&2.85&2.17&8.98&9.42&0.11&7.63
\\   \hline
\texttt{Do} &0.81&5.26&6.60&0.00&9.44&14.21&0.18&15.86&3.73&1.67&1.97&5.88&6.39&0.03&27.98
\\ \hline
\texttt{For} &0.86&6.28&3.19&0.79&-&22.89&0.09&21.08&5.01&3.34&1.87&10.01&10.71&0.08&13.79
\\ \hline
\texttt{If}&1.64&8.43&2.87&0.60&13.49&-&0.24&26.46&7.45&4.80&2.85&9.89&15.11&0.08&6.11
\\ \hline
\texttt{Label} &1.30&8.33&7.86&1.11&5.18&22.85&-&15.17&3.05&2.04&14.62&10.45&4.16&0.09&3.79
\\ \hline
\texttt{Return} &1.13&9.41&3.11&0.49&13.33&27.24&0.24&-&5.59&3.65&2.55&14.91&12.61&0.12&5.61
\\ \hline
\texttt{Case} &0.78&2.84&2.84&0.39&10.27&31.79&0.16&22.40&-&0.46&2.07&7.37&11.69&0.08&6.87
\\ \hline
\texttt{Switch}&1.14&2.72&3.80&0.55&11.07&34.14&0.13&21.86&0.75&-&1.53&8.65&9.02&0.05&4.58
\\ \hline
\texttt{Synch} &0.80&6.57&2.28&0.43&10.21&24.18&0.05&19.77&6.35&2.07&-&9.16&12.16&0.04&5.93
\\ \hline
\texttt{Throw} &2.11&6.57&2.58&0.48&11.87&18.84&0.17&32.28&4.64&3.30&2.74&-&10.08&0.07&4.27
\\ \hline
\texttt{Try}&0.71&7.41&3.02&0.66&11.73&27.75&0.11&23.24&5.63&2.65&2.58&8.99&-&0.09&5.42
\\ \hline
\texttt{TypeDecl} &0.00&4.51&7.52&1.00&10.28&21.05&0.50&17.79&6.02&1.75&2.01&9.27&11.53&-&6.77
\\ \hline
\texttt{While}& 0.72&8.02&3.82&1.96&23.16&19.78&0.12&16.48&6.56&3.09&1.64&6.81&7.80&0.04&-
			\\ \hline
		\end{tabular}
		}
		\caption{Likeliness to replace the statement in the first column for the statement in the top row, in Java}\label{tab:likeliness}
\end{table*}

@@

\section{Statement-level mutations}
Some automatic program repair approaches mutate source code using predefined
mutation operators such as addition, deletion and replacement. In
order to generate patches that make more sense to human developers, it would be
beneficial to analyze how developers mutate source code to fix bugs.

\subsection{Replacements}
Because direct diffs are difficult to identify, we heuristically approximate the
extent to which one statement type appears to be ``replaced'' by another across
this dataset.  To do this, for each modified file, we count the number of
appearances of each statement type in the file pre- and post-commit.  We then
compare the results to see how many of each statement type was removed, and how
many inserted, to roughly characterize the types of replacement that happen at a
per-file level.  For each statement type (as tagged by the Boa infrastructure
for Java), we process the results as follows: 
if number of occurrences of one statement type decreased post-fix and the number of
another type increased, we say that
the first statement type was replaced by the second statement type for that
file.

@@

The numbers in the slots of Table~\ref{tab:likeliness}, represent the percentage rate at which the
elements of the leftmost column were replaced by the elements of the top row. In
the first column we have the list of statements that are being replaced (faulty
statements that were removed), and on the top row we have the list of statements
for which that statement was replaced for (the new statement put instead). For
example, if we take the intersection of row 5 (DoStatement row) and column 2
(AsserStatement column) of the data, we have a 0.81, which means that Do statements were replaced by Assert statements 0.81\% of the times.

@@

\subsubsection{Least common replacements}
On the opposite side of the data, we can see that the least common replacement
was an Assert statement replacing a Type Declaration statement, which has zero appearances. The second least common replacement was
removing a Do statement and insert a Type Declaration statement, with one
appearance in our search, and on third place, removing a Label and insert
a Type Declaration Statement had one appearance in our search.

@@

\begin{figure*}[!t] \centering
	\includegraphics[height=0.3\textheight, width=1\textwidth]{StmtType} \caption{For each kind
	of statement, how many bug fixing revisions add/delete statements of the
same kind?} \label{fig:StmtType} \end{figure*}
\todo{I'm not sure the percentages in Figure~\ref{fig:StmtType} are visible, maybe moving the numbers so they don't overlap with the bars?}
For each bug fixing revision $r$ and each statement kind $k$, we count how many
statements of kind $k$ are there in revision $r$ and $r-1$ ($r-1$ means the
revision before revision $r$), and then compare the two numbers.

@@

For each bug fixing revision, the results discussed above consider both newly
added files and existing files. Since most automatic program repair approaches
do not generate new files to fix bugs, the rest of this section only considers
files that were already in the repositories when developers fixed bugs. Our
analysis shows that there are 46,301,429 of them.

We found that the most common $Replacer$ statement kind (the statement
kind that most commonly replaces others) is the \textbf{If Statement} with
101,366 appearances; and the least common $Replacer$, is the \textbf{Type Declaration
  Statement} with 447 appearances.

The most common $Replacee$ (the most likely statement kind to be replaced
by another statement kind) is the \textbf{Return Statement} with 111,938
appearances; and the least common $Replacee$ is the \textbf{Type Declaration Statement}
with 399 appearances.

\subsection{Applicability to Automatic Error Repair}
This study is directly applicable to automatic software repair. Most common
tools for automatic software repair~\cite{kim2013,weimer2009,legoues2012}
first find a faulty statement using several different fault localization
techniques~\cite{fry2010}, and once they have that faulty statement, they apply
different kinds of edits to it, usually at random. Which makes it equally likely
to replace an statement for another without looking at the kinds of statements
being replaced.

With this study, we are able to make a more accurate guess, when having a known
faulty statement, of which statements more commonly replace that faulty one to build a successful patch based on the data from Table~\ref{tab:likeliness}. Similarly with Figure~\ref{fig:StmtType}, we can have an informed guess of how often programmers insert or delete that particular kind of statement.

For example, if we are fixing a Java program, and we have a faulty statement
that happens to be a For Statement. Then we can look for the most likely statements to replace a For Statement (row 6): If statement (22.89\% likeliness), Return statement (21.08\%), and a While statement (13.79\% of likeliness to replace a For Statement). We also know that there is a 7.9\% of cases in which human programmers deleted this kind of statement, and in 10.6\% of the cases, a For statement was added to patch a bug. This
way we provide guidelines to
the automatic bug repair approaches to consider this on their effort to find a
patch.
